---
title: "Brownlows Bets"
author: "Erika Munoz"
date: "November 15, 2019"
output: html_document: default
 ---

<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 1137 times so far. Where do you think his profits are? Discuss this in a markdown document

# Methods

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 1137 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->


To estimate professor Brownlow's earnings or losses after 1137 bets, the data sample gathered from the bets will be compiled by first setting a new object for birthdays (BDAYS) which will contain each possible day of the year from an interval 1:365. Then using our (BDAYS) object we will compare the sample tests of each bet results to whether or not they obtained a unique sample number < 35 which is the total number of birthdays per bet. After all 1,137 bets there will be a total amount of money for the bets either in positive or negative money total due to earnings or loses. This data will be shown as a 95% confidence interval over a histogram.


#Profit Estimate

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting!-->

```{r}
Bdays = 1:365
Bsamples = vector("numeric")
TotalDollars = 0
saveDollars = vector("numeric")

```

run the simulation 1137 times.



```{r}
for (loop in 1:1000) {
  
  TotalDollars = 0
  classsize = 35
  for(i in 1:1137) {
  Bsamples = sample(Bdays,35, replace= TRUE)
  if (length(unique(Bsamples)) == 35) {
    TotalDollars = TotalDollars -2
  } else{
    TotalDollars = TotalDollars +1
  }  
  }
  
saveDollars[loop] = TotalDollars  
  }

hist(saveDollars)
var(saveDollars)
print(paste("average return =", mean(saveDollars)))
```



This bar graph shows that the average dollars made from professor Brownlow's bets is $505.

# Confidence interval of Brownlows earning

<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows-->

```{r}
varSorted = sort(saveDollars)
LCL = varSorted[25]
UCL = varSorted[975]
print(paste("UCL = ", UCL, "LCL", LCL))
```

For this 95% confidence interval we are shown a UCL of 588 and a LCL of 429. These two limits are taken from our same testing that gave us an average of $505. 



# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

```{r}
for (loop in 1:1000) {
  
  TotalDollars = 0
  classsize = 35
  for(i in 1:1137) {
  Bsamples = sample(Bdays,35, replace= TRUE)
  if (length(unique(Bsamples)) == 35) {
    TotalDollars = TotalDollars -4.40
  } else{
    TotalDollars = TotalDollars +1
  }  
  }
  
saveDollars[loop] = TotalDollars  
  }

hist(saveDollars)
var(saveDollars)
print(paste("average return =", mean(saveDollars)))
```

In this test I ran an R chunk to see how professor Brownlow could break even. I started by betting at -3, then at -100,then -20, then -4.40. I got the average of -0.7206... I finally was able to cut even in my test.

```{r}
for (loop in 1:1000) {
  
  TotalDollars = 0
  classsize = 28
  for(i in 1:1137) {
  Bsamples = sample(Bdays,28, replace= TRUE)
  if (length(unique(Bsamples)) == 28) {
    TotalDollars = TotalDollars -2
  } else{
    TotalDollars = TotalDollars +1
  }  
  }
  
saveDollars[loop] = TotalDollars  
  }

hist(saveDollars)
var(saveDollars)
print(paste("average return =", mean(saveDollars)))
```

For the class size it has to be 28 student because if were try to do 29 students the average return becomes a positive which means that it would be a quart of a student and it is imposible to have a quarter of a student. 
